Big Ideas in Sports Analytics and Statistical Tools for their Investigation

### 2022-11-01 response_to_reviewers.Rmd

We thank the reviewer for this excellent suggestion. We have added a subsection on Elo ratings, with a mathematical definition followed by a calculation example. This example uses the `chessR` package, which ties into and strengthens our discussion of tools in the next chapter.

> There is a large body of literature on Bayesian state-space models in sports outside of the Glickman and Stern paper, mainly in soccer/association football. See for example Koopman and Lit (2015) and the various references in its introduction.

Thank you for this suggestion. We have added a reference to Koopman's paper.

> As mentioned earlier, Section 5 seems sparse on details, and furthermore it seems odd to choose chess as the case study when it has barely been mentioned previously in the article.

Our new subsection on Elo ratings also discusses chess, so now, the reader has already seen a methodological application to chess by the time they get to the case study. We have also added a new subsection to this Section about baseball, in the hopes of fleshing it out.

> I think this section could use a table of some sort, perhaps something that summarizes the CRAN task view at a high level. Another potential would be a figure which shows a typical workflow of sports analytics as explained in the text.

@@

> 12, 1: As noted above, Section 4 feels much weaker than Sections 2 and 3 which are strong.  Perhaps a look at Pythagorean methods as a historical introduction to this topic.

We thank the reviewer for this excellent idea. We have added a new subsection on Expected Winning Percentages models to the beginning of Section 4.

> 12, 23: This is not much of a case study.  Perhaps some additional information on the data available, ways it could be used to calculate game state expected valued, win probability, player ratings….

We have connected the case study in chess to the new subsection on Elo ratings and the subsequent subsection on graphical tools. We hope that these explicit connections to additional content will make this subsection feel more substantial. 
In addition, we have added new subsection to this Section with a different case study in baseball about the evolution of data, tools, and statistical methodologies. We hope that this additional content will help to flesh out the value of these tools.

### 2022-11-01 sports_analytics.Rmd
close #32

[Application programming interfaces](https://en.wikipedia.org/wiki/API) (APIs) are also a common source for data retrieval. 
In chess, the \pkg{chessR} package [@R-chessR] allows R users to download game data from the Chess.com API. 
This type of infrastructure, where one package is the "workhorse" that facilitates common generic data operations, and other packages layer on specific functionality, is common in sports analytics.

Figure \ref{fig:chess} shows a rendering of the starting chess board obtained via the \pkg{chess} package, along with the final position in the game won by one of the authors mentioned earlier in Section \ref{sec:elo} (with data downloaded via the \pkg{chessR} package).

@@

Player tracking data contains coordinates that reveal player movement, and these coordinates are always understood in context relative to reference points on the field, court, ice, or pitch for a particular sport. 
Orienting these points graphically may require drawing a complex set of guidelines that provide that context to readers. 
Thankfully, the \pkg{sportyR} package [@R-sportyR] contains generic playing surfaces for baseball, basketball, curling, American football, ice hockey, soccer, and tennis that can be added to \pkg{ggplot} graphics with a single function call. 
Consider, for example, how much time the surfaces plotted in Figure \ref{fig:sportyR} would save you, and how helpful they would be in contextualizing player tracking data (such as those shown in Figure \ref{fig:epv-demo}).
With the increased availability of player tracking data, this particular tool should see increased usage.

>>>>

[Application programming interfaces](https://en.wikipedia.org/wiki/API) (APIs) are also a common source for data retrieval. 
In chess, the \pkg{chessR} package [@R-chessR] allows R users to download game data from the Chess.com API. 
This type of infrastructure, where one package is the "workhorse" that facilitates common generic data operations, and other packages layer on specific functionality, is common in sports analytics.
See Section \ref{sec:elo} for an example of how the \pkg{chessR} package can be used to compute Elo ratings.

Figure \ref{fig:chess} shows a rendering of the starting chess board obtained via the \pkg{chess} package, along with the final position in the game won by one of the authors mentioned earlier in Section \ref{sec:elo} (with data downloaded via the \pkg{chessR} package). 
We note how the contextual information provided by the chessboard is instrumental in helping the reader contextualize the data. 
(How many of us can visualize PGN directly?)
In Section \ref{sec:graphics}, we outline a collection of graphical tools that provide similar context for different playing surfaces.

@@

Player tracking data contains coordinates that reveal player movement, and these coordinates are always understood in context relative to reference points on the field, court, ice, board, or pitch for a particular sport. 
Orienting these points graphically may require drawing a complex set of guidelines that provide that context to readers. 
Thankfully, the \pkg{sportyR} package [@R-sportyR] contains generic playing surfaces for baseball, basketball, curling, American football, ice hockey, soccer, and tennis that can be added to \pkg{ggplot} graphics with a single function call. 
Consider, for example, how much time the surfaces plotted in Figure \ref{fig:sportyR} would save you, and how helpful they would be in contextualizing player tracking data (such as those shown in Figure \ref{fig:epv-demo}).
With the increased availability of player tracking data, this particular tool should see increased usage.
